### 2024-10-30 paper.qmd
update on paper methods and prior model code
- more samples and chains used, given the computational power available

Despite the importance of RRi analysis, modeling their nonlinear behavior during and after exercise presents significant challenges. Traditional approaches, such as linear regression and time-series analysis, often fail to capture the intricate transitions in RRi, especially during intense physical exertion and recovery. This limitation is due to the non-linear and non-stationary nature of HRV [@gronwald2019effects]. Linear models, which oversimplify these transitions, are prone to misrepresent the actual physiological processes. More advanced non-linear models have begun to address this issue, yet many focus on overall HRV parameters rather than directly modeling the RRi [@gronwald2019non].

@@

Hence, the primary objective of this paper is to present a novel non-linear model that characterizes the fluctuations in RRi during exercise and recovery. This model is specifically designed to reflect the complex, real-time changes in RRi, offering a more precise understanding of the physiological processes underlying cardiovascular adaptation to exercise. By focusing exclusively on RRi, the model can deliver insights directly applicable to athletic training regimens, recovery protocols, and clinical practices aimed at improving cardiovascular health. The proposed model seeks to bridge the gap between existing modeling frameworks and the physiological reality of RRi dynamics during exercise.

@@

The priors were selected based on physiological constraints and are detailed in @eq-priors.

@@

![](figures/prior-model.pdf)

@@

The fitting process utilized four Markov Chain Monte Carlo (MCMC) chains, each consisting of 12,500 iterations with a burn-in period of 5,000 iterations, resulting in a total of 30,000 post-warmup samples.

@@

![](figures/model-constituents.pdf)

@@

![](figures/sobol-sensitivity.pdf)

[**Figure 3**]{#fig-sensitivity}. Sensitivity analysis results illustrating the impact of parameter variability on model predictions. Shaded areas represent 95%, 80%, and 60% CI estimated from Monte Carlo samples.

@@

![](figures/rate-sensitivity.pdf)

@@

Initial exploration of RRi dynamics using two-dimensional density kernel estimation (see [Figure 5](#fig-2d-kernel-density)) indicates a clear drop in RRi around the 5-7 minutes mark, associated with the exercise-induced cardiovascular stress. However, greater variability across individuals in post-exercise recovery can be observed .

![](figures/2d-density-kernel.pdf)

>>>>

Despite the importance of RRi analysis, modeling their nonlinear behavior during and after exercise in one continuous measurement it is not without significant challenges. Traditional approaches, such as linear regression and time-series analysis, often fail to capture the intricate transitions in RRi, especially during intense physical exertion and recovery. This limitation is due to the non-linear and non-stationary nature of HRV [@gronwald2019effects]. Linear models, which oversimplify these transitions, are prone to misrepresent the actual physiological processes. More advanced non-linear models have begun to address this issue, yet many focus on overall HRV parameters rather than directly modeling the RRi [@gronwald2019non].

@@

Hence, the primary objective of this paper is to present a novel non-linear model that characterizes the fluctuations in RRi during exercise and recovery. This model is specifically designed to reflect the complex, real-time changes in RRi, offering meaningful model parameters with the potential to explore a precise understanding of the physiological processes underlying cardiovascular adaptation to exercise. By focusing exclusively on RRi, the model can deliver insights directly applicable to athletic training regimens, recovery protocols, and clinical practices aimed at improving cardiovascular health. The proposed model seeks to bridge the gap between existing modeling frameworks and the physiological reality of RRi dynamics during exercise.

@@

The priors were selected based on physiological constraints and are detailed in @eq-priors, ensuring in this way, identifiability of model parameters by constraining the parameter space to physiologically plausible values and ultimately, improving model convergence.

@@

![](figures/prior-model.png)

@@

The fitting process utilized five Markov Chain Monte Carlo (MCMC) chains, each consisting of 15,000 iterations with a burn-in period of 5,000 iterations, resulting in a total of 50,000 post-warmup samples.

@@

![](figures/model-constituents.png)

@@

![](figures/sobol-sensitivity.png)

[**Figure 3**]{#fig-sensitivity}. Sensitivity analysis results illustrating the impact of parameter variability on model predictions and the percentage of explained variance accounted for each model parameter. Shaded areas represent 95%, 80%, and 60% CI estimated from Monte Carlo samples.

@@

![](figures/rate-sensitivity.png)

@@

Initial exploration of RRi dynamics using two-dimensional density kernel estimation (see [Figure 5](#fig-2d-kernel-density)) indicates a clear drop in RRi around the 5-7 minutes mark, associated with the exercise-induced cardiovascular stress. However, greater variability across individuals in post-exercise recovery can be observed.

![](figures/2d-density-kernel.png)
